Modeling Lake Trophic State: A Data Mining Approach

### 2014-09-11 inst/doc/manuscript.Rmd
updated vignette build from new Rmd

---
title: 'Modeling Lake Trophic State: A Data Mining Approach'
author:
- affiliation: aed
  email: hollister.jeff(at)epa.gov
  footnote: Corresponding author
  name: Jeffrey W. Hollister
- affiliation: aed
  name: W. Bryan Milstead
- affiliation: aed
  name: Betty J. Kreakie
address:
- address: "US Environmental Protection Agency Office of Research and Development National
    Health and Environmental Effects Research Laboratory Atlantic Ecology Division 27
    Tarzwell Drive  Narragansett, RI, 02879, USA  "
  code: aed
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: elsarticle.latex
  html_document: null
  word_document: null
csl: plos.csl
documentclass: elsarticle
layout: review
linenumbers: no
bibliography: LakeTrophicModelling.bib
abstract:  Productivity of lentic ecosystems has been well studied and it is 
  widely accepted that as nutrient inputs increase,  productivity increases and
  lakes transition from low trophic state (e.g. oligotrophic) to higher trophic 
  states  (e.g. eutrophic). These broad trophic state classifications are good 
  predictors of ecosystem health and ecosystem services/disservices (e.g. recreation,
  aesthetics, fisheries, and harmful algal blooms). While the relationship between 
  nutrients and trophic state provides reliable predictions, it requires *in situ* 
  water quality data in order to paramterize the model.  This limits the application 
  of these models to lakes with existing and, more importanly, available water quality
  data.  To expand our ability to predict in lakes without water quality data, we take
  advantage of the availability of a large national lakes water quality database, 
  land use/land cover data, lake morphometry data, other universally available data, 
  and modern data mining approaches to build and assess models of lake tropic state 
  that may be more universally applied.  We use random forests and random forest 
  variable selection to identify variables to be used for predicting trophic state and 
  we compare the performance of two models of trophic state (as determined by 
  chlorophyll a concentration).   The first model estimates trophic state with *in situ* 
  as well as universally available data and the second model uses universally available 
  data only. For each of these models we used three separate trophic state categories, 
  for a total of six models. Overall accuracy for the *in situ* and universal data 
  models ranged from xx% to xx% and xx, xx, and xx described the most variation in 
  trophic state.  For the universal data only models, Overall accuraccy ranged from xx% 
  to xx% and xx, xx, and xx described the most variation in trophic state. Lastly, it 
  is believed that the presence and abundance of cyanobacteria is strongly associated 
  with trophic state.  To test this we examine the association between estimates of 
  cyanobacteria biovolume and the measured and predicted trophic state.  Expanding 
  these preliminary results to include cyanobacteria taxa indicates that cyanobacteria 
  are significantly more likely to be found in highly eutrophic lakes. These results 
  suggest that predictive models of lake trophic state may be improved with additional 
  information on the landscape surrounding lakes and that those models provide 
  additional information on the presence of potentially harmful cyanobacteria taxa. 
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Modeling Lake Trophic State: A Data Mining Approach}
-->

```{r setup, include=FALSE, echo=FALSE}
devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
                         auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
devtools::install_github("wesanderson","jhollist")
library("wesanderson")
library("LakeTrophicModelling")
library("knitr")
library("ggplot2")
library("sp")
library("rgdal")
library("xtable")
library("e1071")
opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
data(hkm2014Data_all)
#Checks for existing cache (from another project)
if(file.exists("cache")){
  for(i in gsub(".rdb","",list.files("cache",".rdb",full.names=T))){
    lazyLoad(i)
  }
}
```

@@

##Data and Study Area
We utilize four primary sources of data for this study,the National Lakes Assessment (NLA), the National Lake Cover Dataset (NLCD), modeled lake morphometery,  and estimated cyanobacteria biovolumes [@usepa2009national;@homer2004development;@xian2009updating;@hollister2010volume;@hollister_predicting_2011;@lakemorpho2014;@beaulieu2013nutrients].  All datasets are national in scale and provide a unique snapshot view of the condition of lakes in the United States'.

@@

```{r Table1, results='asis', echo=FALSE}
translation<-c("Percent Impervious","Percent Water","Percent Ice/Snow","Percent Developed Open Space",
               "Percent Low Intensity Development","Percent Medium Intensity Development",
               "Percent High Intensity Development","Percent Barren","Percent Decidous Forest",
               "Percent Evergreen Forest","Percent Mixed Forest","Precent Shrub/Scrub","Percent Grassland",
               "Percent Pasture", "Percent Cropland", "Percent Woody Wetland", 
               "Percent Herbaceuos Wetland","Longitude","Latitude","Lake Surface Area", "Lake Perimeter",
               "Shoreline Development Index","Date Samples Collected", "Ecoregion", "Watershed Area",
               "Maximum Depth","Elevation","Dissolved Oxygen","pH","Conductivity",
               "Acid Neutralizing Capacity","Turbidity","Total Organic Carbon","Dissolved Organic Carbon",
               "Ammonium","Nitrate/Nitrite", "Total Nitrogen","Total Phosphorus","Chloride","Nitrate",
               "Sulfate","Calcium","Magnesium","Sodium","Potassium","Color","Silica","Hydrogen Ions",
               "Hydroxide","Calculate Ammonium","Cation Sum","Anion Sum","Anion Deficit","Base Cation Sum",
               "Ion Balance","Estimate Organic Anions","Calculated Conductivity",
               "D-H-O Calculated Conductivity","Mean Profile Water Temperature","Growing Degree Days",
               "Maximum Lake Length","Maximum Lake Width","Mean Lake Width","Fetch from North", 
               "Fetch form Northeast","Fetch from East", "Fetch from Southeast",
               "Estimated Maximum Lake Depth","Estimated Lake Volume","Estimated Mean Lake Depth",
               "Nitrogen:Phophorus Ratio")
wq<-!predictors_all%in%predictors_gis
type<-vector("character",length(predictors_all))
type[wq]<-"Water Quality"
type[!wq]<-"GIS"
data_def<-data.frame(Variables=predictors_all,Description=translation,Type=type)
# This is an unbelievably silly table... It is life sized!
kable(data_def)
#dfToImage(data_def,style="css/texttable.css",file="figure/data_def_table.jpg")
```

```{r results='asis', echo=FALSE}
state<-map_data('state')
lakes_alb<-data.frame(hkm2014Data[["AlbersX"]],hkm2014Data[["AlbersY"]])
p4s<-"+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
ll<-"+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
lakes_alb_sp<-SpatialPoints(coordinates(lakes_alb),proj4string=CRS(p4s))
lakes_dd<-spTransform(lakes_alb_sp,CRS=CRS(ll))
lakes_dd<-data.frame(coordinates(lakes_dd))
mycolor<-wes.palette(5,"Zissou2")
mycolor<-c(mycolor[1],mycolor[2],mycolor[4])
names(lakes_dd)<-c("long","lat",c())
nlaMap(state,lakes_dd,mycolor)
#ggsave("figure/nla_map.jpg",nlaMap(state,lakes_dd,mycolor),
#       width=18,units="in",dpi=600)
```
**Predicting Trophic State with Random Forests**

@@

```{r trophicStateTable, results='asis', echo=FALSE, fig.cap="Table 2. Chlorphyll a based trophic state cut-offs"}
ts_4<-c("oligo","meso","eu","hyper")
ts_3<-c("oligo","meso/eu","meso/eu","hyper")
ts_2<-c("oligo/meso","oligo/meso","eu/hyper","eu/hyper")
co<-c("<= 0.2",">2-7",">7-30",">30")
xdf<-data.frame(ts_4,ts_3,ts_2,co)
names(xdf)<-c("Trophic State (4)","Trophic State (3)","Trophic State (2)","Cut-off")
kable(xdf)
#dfToImage(xdf,style="css/texttable.css",file="figure/ts_classes_table.jpg")
```

@@

```{r Figure_Model1,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
#dfToImage(sumTable(all_ts4[1:100]),file="figure/all_ts4_table.jpg",
#         style="css/sumtable.css",width=8,include.rownames=FALSE)
#dfToImage(formatC(round(all_ts4_rf$confusion,2)),file="figure/all_ts4_confusion.jpg",
#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
kable(sumTable(all_ts4[1:100]),row.names=F)
kable(formatC(round(all_ts4_rf$confusion,2)),row.names=F)

all_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts4_vars]
importancePlot(all_ts4_rf,'acc',size=10,aes(colour=all_ts4_color))
importancePlot(all_ts4_rf,'gini',size=10,aes(colour=all_ts4_color))
```

@@

@@

```{r Figure_Model2,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
#dfToImage(sumTable(all_ts3[1:100]),file="figure/all_ts3_table.jpg",
#          style="css/sumtable.css",width=8,include.rownames=FALSE)
#dfToImage(formatC(round(all_ts3_rf$confusion,2)),file="figure/all_ts3_confusion.jpg",
#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
kable(sumTable(all_ts3[1:100]),row.names=F)
kable(formatC(round(all_ts3_rf$confusion,2)),row.names=F)

all_ts3_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts3_vars]
importancePlot(all_ts3_rf,'acc',size=10,aes(colour=all_ts3_color))
importancePlot(all_ts3_rf,'gini',size=10,aes(colour=all_ts3_color))
```

@@

@@

```{r Figure_Model3,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
#dfToImage(sumTable(all_ts2[1:100]),file="figure/all_ts2_table.jpg",
#          style="css/sumtable.css",width=8,include.rownames=FALSE)
#dfToImage(formatC(round(all_ts2_rf$confusion,2)),file="figure/all_ts2_confusion.jpg",
#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
kable(sumTable(all_ts2[1:100]),row.names=F)
kable(formatC(round(all_ts2_rf$confusion,2)),row.names=F)

all_ts2_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts2_vars]
importancePlot(all_ts2_rf,'acc',size=10,aes(colour=all_ts2_color))
importancePlot(all_ts2_rf,'gini',size=10,aes(colour=all_ts2_color))
```

@@

@@

```{r Figure_Model4,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8,fig.cap="Figure 1"}
#dfToImage(sumTable(gis_ts4[1:100]),file="figure/gis_ts4_table.jpg",
#          style="css/sumtable.css",width=8,include.rownames=FALSE)

#dfToImage(formatC(round(gis_ts4_rf$confusion,2)),file="figure/gis_ts4_confusion.jpg",
#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
kable(sumTable(gis_ts4[1:100]),row.names=F)
kable(formatC(round(gis_ts4_rf$confusion,2)),row.names=F)

gis_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts4_vars]
importancePlot(gis_ts4_rf,'acc',size=10,aes(colour=gis_ts4_color))
importancePlot(gis_ts4_rf,'gini',size=10,aes(colour=gis_ts4_color))
```

@@

@@

```{r Figure_Model5,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
#dfToImage(sumTable(gis_ts3[1:100]),file="figure/gis_ts3_table.jpg",
#          style="css/sumtable.css",width=8,include.rownames=FALSE)
#dfToImage(formatC(round(gis_ts3_rf$confusion,2)),file="figure/gis_ts3_confusion.jpg",
#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
kable(sumTable(gis_ts3[1:100]),row.names=F)
kable(formatC(round(gis_ts3_rf$confusion,2)),row.names=F)

gis_ts3_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts3_vars]
importancePlot(gis_ts3_rf,'acc',size=10,aes(colour=gis_ts3_color))
importancePlot(gis_ts3_rf,'gini',size=10,aes(colour=gis_ts3_color))
```

@@

@@

```{r Figure_Model6,results="asis",echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
#dfToImage(sumTable(gis_ts2[1:100]),file="figure/gis_ts2_table.jpg",
#          style="css/sumtable.css",width=8,include.rownames=FALSE)
#dfToImage(formatC(round(gis_ts2_rf$confusion,2)),file="figure/gis_ts2_confusion.jpg",
#          style="css/confusiontable.css",width=8,include.rownames=TRUE)
kable(sumTable(gis_ts2[1:100]),row.names=F)
kable(formatC(round(gis_ts2_rf$confusion,2)),row.names=F)

gis_ts2_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts2_vars]
importancePlot(gis_ts2_rf,'acc',size=10,aes(colour=gis_ts2_color))
importancePlot(gis_ts2_rf,'gini',size=10,aes(colour=gis_ts2_color))
```

@@

@@

```{r ts_4_biov, echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
plotCdf(hkm2014Data$TS_CHLA_4,hkm2014Data$sumBioV+1,
                                     cdf_colors=zissou[6:9],y='Percent',
                                     x='Log10(Cyanobacteria Biovolume)',
                                     title=expression(paste('CDF for Chlorophyll ', 
                                     italic("a"),' Trophic States (4 Categories)')),
                                     color="Trophic State\nCategories")
```

```{r ts_3_biov, echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
plotCdf(hkm2014Data$TS_CHLA_3,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:8],
       y='Percent',x='Log10(Cyanobacteria Biovolume)',
       title=expression(paste('CDF for Chlorophyll ', italic("a"),' Trophic States (3 Categories)')),
       color="Trophic State\nCategories")
```

```{r ts_2_biov, echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
plotCdf(hkm2014Data$TS_CHLA_2,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:7],
       y='Percent',x='Log10(Cyanobacteria Biovolume)',
       title=expression(paste('CDF for Chlorophyll ', italic("a"),' Trophic States (2 Categories)')),
       color="Trophic State\nCategories")
```

```{r scatterplot,echo=FALSE, dev="jpeg", fig.width=8, fig.height=8}
scp_df<-data.frame(chla=hkm2014Data[["CHLA"]],biovp1=hkm2014Data[["sumBioV"]]+1)
scatterPlot(scp_df,xvar="chla",yvar="biovp1",zissou[8],zissou[7],zissou[6],
             title=expression(paste("Chlorophyll ", italic("a")," and Cyanobacteria Relationship")),
             x=expression(paste('Log10(Chl ', italic("a"),')')),
             y="Log10(Cyanobaterial Biovolumes + 1)")
```

>>>>

---
title: 'Modeling Lake Trophic State: A Data Mining Approach'
author:
- affiliation: aed
  email: hollister.jeff(at)epa.gov
  footnote: Corresponding author
  name: Jeffrey W. Hollister
- affiliation: aed
  name: W. Bryan Milstead
- affiliation: aed
  name: Betty J. Kreakie
address:
- address: "US Environmental Protection Agency Office of Research and Development National
    Health and Environmental Effects Research Laboratory Atlantic Ecology Division 27
    Tarzwell Drive  Narragansett, RI, 02879, USA  "
  code: aed
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: elsarticle.latex
  html_document: null
  word_document: null
fontsize: 11pt
csl: plos.csl
documentclass: elsarticle
layout: review
linenumbers: yes
bibliography: LakeTrophicModelling.bib
abstract:  Productivity of lentic ecosystems has been well studied and it is 
  widely accepted that as nutrient inputs increase,  productivity increases and
  lakes transition from low trophic state (e.g. oligotrophic) to higher trophic 
  states  (e.g. eutrophic). These broad trophic state classifications are good 
  predictors of ecosystem health and ecosystem services/disservices (e.g. recreation,
  aesthetics, fisheries, and harmful algal blooms). While the relationship between 
  nutrients and trophic state provides reliable predictions, it requires *in situ* 
  water quality data in order to paramterize the model.  This limits the application 
  of these models to lakes with existing and, more importanly, available water quality
  data.  To expand our ability to predict in lakes without water quality data, we take
  advantage of the availability of a large national lakes water quality database, 
  land use/land cover data, lake morphometry data, other universally available data, 
  and modern data mining approaches to build and assess models of lake tropic state 
  that may be more universally applied.  We use random forests and random forest 
  variable selection to identify variables to be used for predicting trophic state and 
  we compare the performance of two models of trophic state (as determined by 
  chlorophyll a concentration).   The first model estimates trophic state with *in situ* 
  as well as universally available data and the second model uses universally available 
  data only. For each of these models we used three separate trophic state categories, 
  for a total of six models. Overall accuracy for the *in situ* and universal data 
  models ranged from xx% to xx% and xx, xx, and xx described the most variation in 
  trophic state.  For the universal data only models, Overall accuraccy ranged from xx% 
  to xx% and xx, xx, and xx described the most variation in trophic state. Lastly, it 
  is believed that the presence and abundance of cyanobacteria is strongly associated 
  with trophic state.  To test this we examine the association between estimates of 
  cyanobacteria biovolume and the measured and predicted trophic state.  Expanding 
  these preliminary results to include cyanobacteria taxa indicates that cyanobacteria 
  are significantly more likely to be found in highly eutrophic lakes. These results 
  suggest that predictive models of lake trophic state may be improved with additional 
  information on the landscape surrounding lakes and that those models provide 
  additional information on the presence of potentially harmful cyanobacteria taxa. 
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Modeling Lake Trophic State: A Data Mining Approach}

-->

```{r setup, include=FALSE, echo=FALSE}
#devtools::install_github("LakeTrophicModelling","USEPA",quick=TRUE,
#                         auth_token="8742a8c1c2207393659a1de1bd53527fa56ea0d7")
#devtools::install_github("wesanderson","jhollist")
library("wesanderson")
library("LakeTrophicModelling")
library("knitr")
library("ggplot2")
library("sp")
library("rgdal")
library("xtable")
library("e1071")
opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=4.3)
data(hkm2014Data_all)

#Checks for existing cache (from another project)
if(file.exists("cache")){
  for(i in gsub(".rdb","",list.files("cache",".rdb",full.names=T))){
    lazyLoad(i)
  }
}

# Table Captions from @DeanK on http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t

knit_hooks$set(tab.cap = function(before, options, envir) {
  if(!before)
    paste('\n\n:', options$tab.cap, sep='') 
})
default_output_hook = knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  if (is.null(options$tab.cap) == F)  
    x
  else
    default_output_hook(x,options)
})
```

@@

##Data and Study Area

We utilize four primary sources of data for this study,the National Lakes Assessment (NLA), the National Lake Cover Dataset (NLCD), modeled lake morphometery,  and estimated cyanobacteria biovolumes [@usepa2009national;@homer2004development;@xian2009updating;@hollister2010volume;@hollister_predicting_2011;@lakemorpho2014;@beaulieu2013nutrients].  All datasets are national in scale and provide a unique snapshot view of the condition of lakes in the United States'.

@@

```{r Variables, results='asis', eval=FALSE, echo=FALSE}
#tab.cap="Definition of all variables considered"
translation<-c("Percent Impervious","Percent Water","Percent Ice/Snow","Percent Developed Open Space","Percent Low Intensity Development","Percent Medium Intensity Development", "Percent High Intensity Development","Percent Barren","Percent Decidous Forest","Percent Evergreen Forest","Percent Mixed Forest","Precent Shrub/Scrub","Percent Grassland", "Percent Pasture", "Percent Cropland", "Percent Woody Wetland", "Percent Herbaceuos Wetland","Longitude","Latitude","Lake Surface Area", "Lake Perimeter","Shoreline Development Index","Date Samples Collected", "Ecoregion", "Watershed Area","Maximum Depth","Elevation","Dissolved Oxygen","pH","Conductivity","Acid Neutralizing Capacity","Turbidity","Total Organic Carbon","Dissolved Organic Carbon","Ammonium","Nitrate/Nitrite", "Total Nitrogen","Total Phosphorus","Chloride","Nitrate","Sulfate","Calcium","Magnesium","Sodium","Potassium","Color","Silica","Hydrogen Ions","Hydroxide","Calculate Ammonium","Cation Sum","Anion Sum","Anion Deficit","Base Cation Sum","Ion Balance","Estimate Organic Anions","Calculated Conductivity","D-H-O Calculated Conductivity","Mean Profile Water Temperature","Growing Degree Days","Maximum Lake Length","Maximum Lake Width","Mean Lake Width","Fetch from North", "Fetch form Northeast","Fetch from East", "Fetch from Southeast","Estimated Maximum Lake Depth","Estimated Lake Volume","Estimated Mean Lake Depth","Nitrogen:Phophorus Ratio")
wq<-!predictors_all%in%predictors_gis
type<-vector("character",length(predictors_all))
type[wq]<-"Water Quality"
type[!wq]<-"GIS"
data_def<-data.frame(Variables=predictors_all,Description=translation,Type=type)
# This is an unbelievably silly table... It is life sized!
kable(data_def)
#dfToImage(data_def,style="css/texttable.css",file="figure/data_def_table.jpg")
```

```{r results='asis', echo=FALSE, fig.cap="Map of the distribution of National Lakes Assesment Sampling locations"}
state<-map_data('state')
lakes_alb<-data.frame(hkm2014Data[["AlbersX"]],hkm2014Data[["AlbersY"]])
p4s<-"+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
ll<-"+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
lakes_alb_sp<-SpatialPoints(coordinates(lakes_alb),proj4string=CRS(p4s))
lakes_dd<-spTransform(lakes_alb_sp,CRS=CRS(ll))
lakes_dd<-data.frame(coordinates(lakes_dd))
mycolor<-wes.palette(5,"Zissou2")
mycolor<-c(mycolor[1],mycolor[2],mycolor[4])
names(lakes_dd)<-c("long","lat",c())
nlaMap(state,lakes_dd,mycolor)
```

**Predicting Trophic State with Random Forests**

@@

```{r trophicStateTable, results='asis', echo=FALSE, tab.cap="Chlorphyll a based trophic state cut-offs"}
ts_4<-c("oligo","meso","eu","hyper")
ts_3<-c("oligo","meso/eu","meso/eu","hyper")
ts_2<-c("oligo/meso","oligo/meso","eu/hyper","eu/hyper")
co<-c("<= 0.2",">2-7",">7-30",">30")
xdf<-data.frame(ts_4,ts_3,ts_2,co)
names(xdf)<-c("Trophic State (4)","Trophic State (3)","Trophic State (2)","Cut-off")
kable(xdf)
```

@@

@@

```{r VarSel_Model1,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 1"}
kable(sumTable(all_ts4[1:100]),row.names=F)
```

```{r Confusion_Model1,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 1"}
kable(formatC(round(all_ts4_rf$confusion,2)),row.names=F)
```

```{r Importance_Model1,results="asis",echo=FALSE, fig.cap="Importance plot for Model 1"}
all_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts4_vars]
importancePlot(all_ts4_rf,'gini',size=10,aes(colour=all_ts4_color))
```

@@

@@

```{r VarSel_Model2,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 2"}
kable(sumTable(all_ts3[1:100]),row.names=F)
```

```{r Confusion_Model2,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 2"}
kable(formatC(round(all_ts3_rf$confusion,2)),row.names=F)
```

```{r Importance_Model2,results="asis",echo=FALSE, fig.cap="Importance plot for Model 2"}
all_ts3_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts3_vars]
importancePlot(all_ts3_rf,'gini',size=10,aes(colour=all_ts3_color))
```

@@

@@

```{r VarSel_Model3,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 3"}
kable(sumTable(all_ts2[1:100]),row.names=F)
```

```{r Confusion_Model3,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 3"}
kable(formatC(round(all_ts2_rf$confusion,2)),row.names=F)
```

```{r Importance_Model3,results="asis",echo=FALSE, fig.cap="Importance plot for Model 3"}
all_ts2_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%all_ts2_vars]
importancePlot(all_ts2_rf,'gini',size=10,aes(colour=all_ts2_color))
```

@@

@@

```{r VarSel_Model4,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 4" }
kable(sumTable(gis_ts4[1:100]),row.names=F)
```

```{r Confusion_Model4,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 4" }
kable(formatC(round(gis_ts4_rf$confusion,2)),row.names=F)
```

```{r Importance_Model4,results="asis",echo=FALSE, tab.cap="Importance plot for Model 3"}
gis_ts4_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts4_vars]
importancePlot(gis_ts4_rf,'gini',size=10,aes(colour=gis_ts4_color))
```

@@

@@

```{r VarSel_Model5,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 5"}
kable(sumTable(gis_ts3[1:100]),row.names=F)
```

```{r Conufsion_Model5,results="asis",echo=FALSE, tab.cap="Random Forest confusion matrix for Model 5"}
kable(formatC(round(gis_ts3_rf$confusion,2)),row.names=F)
```

```{r Importance_Model5,results="asis",echo=FALSE, fig.cap="Importance plot for Model 5"}
gis_ts3_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts3_vars]
importancePlot(gis_ts3_rf,'gini',size=10,aes(colour=gis_ts3_color))
```

@@

@@

```{r VarSel_Model6,results="asis",echo=FALSE, tab.cap="Variable selection results for Model 6"}
kable(sumTable(gis_ts2[1:100]),row.names=F)
```

```{r Confusion_Model6,results="asis",echo=FALSE, tab.cap="Random forest confusion matrix for Model 6"}
kable(formatC(round(gis_ts2_rf$confusion,2)),row.names=F)
```

```{r Importance_Model6,results="asis",echo=FALSE, fig.cap="Importance plot for Model 6"}
gis_ts2_color<-col_lu[["hexcode"]][col_lu[["variables"]]%in%gis_ts2_vars]
importancePlot(gis_ts2_rf,'gini',size=10,aes(colour=gis_ts2_color))
```

@@

```{r ts_4_biov, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 4 trophic state classes"}
plotCdf(hkm2014Data$TS_CHLA_4,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:9],
        y='Percent',x='Log10(Cyanobacteria Biovolume)',
        title=expression(paste('CDF for Chlorophyll ', 
                                italic("a"),' Trophic States (4 Categories)')),
        color="Trophic State\nCategories")
```

```{r ts_3_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 3 trophic state classes" }
plotCdf(hkm2014Data$TS_CHLA_3,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:8],
       y='Percent',x='Log10(Cyanobacteria Biovolume)', 
       title=expression(paste('CDF for Chlorophyll ', 
                              italic("a"),' Trophic States (3 Categories)')),
       color="Trophic State\nCategories")
```

```{r ts_2_biov_cdf, echo=FALSE, fig.width=7, fig.cap="Cumulative distribution function of cyanobacetria biovolume for 2 trophic state classes"}
plotCdf(hkm2014Data$TS_CHLA_2,hkm2014Data$sumBioV+1,cdf_colors=zissou[6:7],
       y='Percent',x='Log10(Cyanobacteria Biovolume)',
       title=expression(paste('CDF for Chlorophyll ', 
                              italic("a"),' Trophic States (2 Categories)')),
       color="Trophic State\nCategories")
```

```{r scatterplot,echo=FALSE, fig.width=7, fig.cap="Cholorphyll *a* and cyanobacteria biovolume scatterplot"}
scp_df<-data.frame(chla=hkm2014Data[["CHLA"]],biovp1=hkm2014Data[["sumBioV"]]+1)
scatterPlot(scp_df,xvar="chla",yvar="biovp1",zissou[8],zissou[7],zissou[6],
             title=expression(paste("Chlorophyll ", 
                                    italic("a")," and Cyanobacteria Relationship")),
             x=expression(paste('Log10(Chl ', italic("a"),')')),
             y="Log10(Cyanobaterial Biovolumes + 1)")
```

\newpage
